### 2021-01-01 ch/3-target-labels.tex

\chapter{Tuning Fairness by Balancing Target Labels}\label{ch:target-labels}