### 2021-01-01 ch/4-nifr.tex

\chapter{Null-sampling for Interpretable and Fair Representations}\label{ch:nifr}