### 2021-01-01 ch/5-zsf.tex

\chapter{Zero-shot Fairness with Invisible Demographics}\label{ch:zsf}