### 2021-01-06 ch/3-target-labels.tex
Finish first draft of chapter 3 (now 4)

\chapter{Tuning Fairness by Balancing Target Labels}\label{ch:target-labels}
This chapter is predominantly concerned with label bias.
The main idea is that we make use of pseudo labels to implicitly learn from a \emph{balanced} dataset,
in which \(y\perp s\) holds, and which thus satisfies \acf{DP}.
This falls under the area of \emph{fair classifiers} discussed in \secref{fair-classifiers}

We can interpret the contributions of this chapter in two ways.
The first corresponds to the definition-centric view of dataset bias,
and the second to the groundtruth-centric view.
\begin{enumerate}
  \item
    We can say that the classifier should satisfy demographic parity in its predictions,
    and learning from a balanced training set is just one particular way to achieve this.
    In this view, the pseudo labels have no deeper meaning and are just a computational trick.
  \item
    We can see the training set as a corrupted version of a true dataset, which is balanced (\(y\perp s\)),
    and so, by learning from these pseudo labels, we are simply approximating the true dataset.
    However, we do not actually have access to the true dataset; we only know that it is balanced.
    In order to evaluate the trained model, we compute fairness metrics with respect to \ac{DP}.
\end{enumerate}
Within the chapter, we sometimes jump between these two views.

While the main focus is on label bias,
the experiments are performed on real-world fairness datasets,
which also display a significant amount of sample bias (disadvantaged groups are underrepresented).
Furthermore, in addition to the result for \ac{DP}, we also show that the proposed scheme improves \acf{EOpp}.

In order to construct the pseudo labels,
we use side information about summary statistics for an unbalanced training set.
This allows us to target a specific balanced set, instead of just any balanced set.
In other words, rather than just enforcing \ac{DP}, the method gives control over the target rates \(P(\hat{y}=1|s)\).