### 2021-01-12 ch/5-zsf.tex

\chapter{Zero-shot Fairness with Invisible Demographics}\label{ch:zsf}
A shortcoming of the approach from the previous chapter
is its reliance on a context/pre-training set which has \(s\) labels.
While it is usually easier to collect data without \(y\) labels and with \(s\) labels,
it is even easier to collect data without any label.