### 2021-02-23 ch/4-nifr.tex
Lots of improvements

\chapter{Content}\label{ch:content}
\section{Tuning Fairness by Balancing Target Labels}\label{sec:target-labels}
This chapter is predominantly concerned with label bias.
The main idea is that we make use of pseudo labels to implicitly learn from a \emph{balanced} dataset,
in which \(y\perp s\) holds, and which thus satisfies \acf{DP}.
This falls under the area of \emph{fair classifiers} discussed in \secref{fair-classifiers}

We can interpret the contributions of this chapter in two ways.
The first corresponds to the definition-centric view of dataset bias,
and the second to the groundtruth-centric view.
\begin{enumerate}
  \item
    We can say that the classifier should satisfy demographic parity in its predictions,
    and learning from a balanced training set is just one particular way to achieve this.
    In this view, the pseudo labels have no deeper meaning and are just a computational trick.
  \item
    We can see the training set as a corrupted version of a true dataset, which is balanced (\(y\perp s\)),
    and so, by learning from these pseudo labels, we are simply approximating the true dataset.
    However, we do not actually have access to the true dataset; we only know that it is balanced.
    In order to evaluate the trained model, we compute fairness metrics with respect to \ac{DP}.
\end{enumerate}
Within the chapter, we sometimes jump between these two views.

While the main focus is on label bias,
the experiments are performed on real-world fairness datasets,
which also display a significant amount of sample bias (disadvantaged groups are underrepresented).
Furthermore, in addition to the result for \ac{DP}, we also show that the proposed scheme improves \acf{EOpp}.

In order to construct the pseudo labels,
we use side information about summary statistics for an unbalanced training set.
This allows us to target a specific balanced set, instead of just any balanced set.
In other words, rather than just enforcing \ac{DP}, the method gives control over the target rates \(P(\hat{y}=1|s)\).

\section{Null-sampling for Interpretable and Fair Representations}\label{sec:nifr}
In the setting from the previous chapter,
labels were untrustworthy because they had been flipped;
a phenomenon we referred to as \emph{label bias}.
However, flipped labels is not the only way that labels can become untrustworthy.
Another way is \emph{sampling bias}, which is the subject of this and the next chapter.

As an example, consider a dataset where the task is to distinguish smiling from non-smiling faces.
Say, we initially have a very diverse and balanced training set,
but then, from the set of smiling faces, we remove nearly all samples where the person does not have red hair,
and from the set of non-smiling faces, we remove nearly all samples with non-black hair.
The result is that even though the labels are still correct, they do not point reliably to the right target anymore.
As hair color is easier to recognize with a \ac{CNN}, the labels have effectively been turned into hair color labels.
In other words, hair color has become a \emph{spurious attribute}.
In the following, we denote the spurious attribute with \(s\),
as it takes a role that is very similar to that of the sensitive attribute that was also denoted by \(s\).

The method, proposed in the previous chapter,
is not able to deal with such a dataset bias as we can easily see:
Say, \emph{smiling} corresponds to \(y=1\) and \emph{not smiling} to \(y=0\);
furthermore, let red hair correspond to \(s=1\), black hair to \(s=0\), and all other hair colors to \(s=2\).
Then, the problem with the described dataset is, that it mostly consists of samples with \(y=0\wedge s=0\)
and those with \(y=1\wedge s=1\).
If we call \(P(y=1|s=s')\) the acceptance rate,
then the problem can be described as one of very different acceptance rates in the hair color groups given by \(s\).
This is the problem tackled in the previous chapter,
and yet, if we were to equalize the acceptance rates with the method there,
the result would be very incorrect.
The issue is that we would treat the labels as incorrect, when in truth, they are correct.

To deal with sampling bias, a different approach is needed.
Indeed, the problem, as posed, is not solvable in the general case.
To make headway with this problem, we introduce the concept of a \emph{context set} (or \emph{pre-training set}).
This set is not subject to the sampling bias,
but is unlabeled (with respect to $y$) and so does not by itself suffice for training.
However, this set does have labels for the spurious attribute \(s\).
This allows us to learn an \emph{invariant representation},
\ie, a representation of the input features \(x\) which is invariant to the spurious attribute.
This kind of representation is equivalent to a fair representation -- as described in \chapref{ch:related-work} --
which is invariant to a sensitive attribute.
With the invariant representation of the training set,
a classifier can then be trained to accurately predict the class label \(y\).

A parallel to the previous chapter is that the method makes use of side information
(in this case the pre-training/context set)
in order to overcome the bias in the training set.

\section{Zero-shot Fairness with Invisible Demographics}\label{sec:zsf}
A shortcoming of the approach from the previous chapter
is its reliance on a context/pre-training set which has \(s\) labels.
As discussed, it is necessary to make use of \emph{some} kind of side information,
but perhaps we can relax some requirements.
In particular, while it is already easier to collect data without \(y\) labels (but with \(s\) labels),
it is even easier to collect data without \emph{any} label.
So, requiring only an unlabeled context set would improve the applicability of the method.
This chapter presents an approach based on that idea.

The setting is very similar to the previous chapter:
the training set suffers from severe sampling bias, but we have access to a (mostly) unbiased context set.
The change is that the context set may be completely unlabeled,
but in exchange, we have some stronger requirements for the training set.
Namely, the holes left by the sampling bias may not be so numerous
as to make \(s\) and \(y\) completely indistinguishable.
In the previous chapter, the example of colored MNIST had a training set
where there was a strict one-to-one mapping of color and digit;
but this kind of blending of \(s\) and \(y\) into one is not the focus of this chapter.
The focus is instead on a setting where the training set lacks certain combinations of \(s\) and \(y\),
which results in poor predictions for these combinations on the test set (or deployment setting)
where those combinations do occur.
We refer to these missing combinations as \emph{invisible demographics}.
The value of \(s\) acts as a spurious attribute and should be ignored in order to make accurate predictions.

As in the previous chapter, the first step is to train a neural network to produce an invariant representation.
The second step is then to train a classifier on that.
The invariant representation is trained by doing distribution matching between the training set and the context set.
Finally, the training set is transformed into the invariant representation,
to serve as training set for the classifier.