### 2021-03-04 FrontBackmatter/Declaration.tex
Make it compile with the 3rd paper

%*******************************************************
% Declaration
%*******************************************************
\pdfbookmark[0]{Declaration}{declaration}
\chapter*{Declaration}
\thispagestyle{empty}
Put your declaration here.
\bigskip

\noindent\textit{\myLocation, \myTime}

\smallskip

\begin{flushright}
    \begin{tabular}{p{5cm}}
        \\ \hline
        \centering\myName \\
    \end{tabular}
\end{flushright}