### 2021-03-08 FrontBackmatter/Declaration.tex
Lots of small improvements

%*******************************************************
% Declaration
%*******************************************************
\pdfbookmark[0]{Declaration}{declaration}
\chapter*{Declaration}
\thispagestyle{empty}
I hereby declare that this thesis has not been and will not be, submitted in whole or in part to another University for the award of any other degree.
Except where indicated by specific stated in the text, this thesis was composed by myself and the work contained therein in my own.
\bigskip

\noindent\textit{\myLocation, \myTime}

\smallskip

\begin{flushright}
    \begin{tabular}{p{5cm}}
        \\ \hline
        \centering\myName \\
    \end{tabular}
\end{flushright}