### 2021-03-16 FrontBackmatter/Acknowledgments.tex
Try to write acknowledgments

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

% \begin{flushright}{\slshape
%     We have seen that computer programming is an art, \\
%     because it applies accumulated knowledge to the world, \\
%     because it requires skill and ingenuity, and especially \\
%     because it produces objects of beauty.} \\ \medskip
%     --- \defcitealias{knuth:1974}{Donald E. Knuth}\citetalias{knuth:1974} \citep{knuth:1974}
% \end{flushright}
% \bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
This thesis would not have been possible without...

I would like to thank first and foremost my supervisor Novi.
He has taught me a whole lot in the time I was his PhD student and always pushed me to challenge myself.

I would like to thank my wife Chao who always supported me perfectly
and encouraged me when I was feeling down.
I would not have made it through the PhD without her.

I would like to thank my office mates Oliver and Myles
who were always there to talk about machine learning and other things as well.
Working together with them was an absolute joy.

Finally, I thank my family who have always believed in me.

\endgroup