### 2021-07-23 FrontBackmatter/Acknowledgments.tex
Clean up

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgements}{acknowledgments}

% \begin{flushright}{\slshape
%     We have seen that computer programming is an art, \\
%     because it applies accumulated knowledge to the world, \\
%     because it requires skill and ingenuity, and especially \\
%     because it produces objects of beauty.} \\ \medskip
%     --- \defcitealias{knuth:1974}{Donald E. Knuth}\citetalias{knuth:1974} \citep{knuth:1974}
% \end{flushright}
% \bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgements}
% Joe's acknowledgements
% I would like to start by thanking my family for all their constant encouragement: my parents Ged and Marian, my sister Dee, my niece Hannah, and my nephews Alfie and Alex. Thank-you to my wonderful partner Clemy who has been amazing and supportive of me. Thanks to everyone I have shared an office with, and last but not least, thanks to my supervisors Novi and David.
I would like to thank first and foremost my supervisor Novi.
He has taught me a whole lot in the time I was his PhD student and always pushed me to challenge myself.

I would like to thank my wife Chao who always supported me perfectly
and encouraged me when I was feeling down.
I would not have made it through the PhD without her.

I would like to thank my office mates Oliver and Myles
who were always there to talk about machine learning and other things as well.
Working together with them was an absolute joy.

Finally, I thank my family who have always believed in me.

\endgroup