### 2021-09-15 FrontBackmatter/Acknowledgments.tex
Clean up

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgements}{acknowledgments}

% \begin{flushright}{\slshape
%     We have seen that computer programming is an art, \\
%     because it applies accumulated knowledge to the world, \\
%     because it requires skill and ingenuity, and especially \\
%     because it produces objects of beauty.} \\ \medskip
%     --- \defcitealias{knuth:1974}{Donald E. Knuth}\citetalias{knuth:1974} \citep{knuth:1974}
% \end{flushright}
% \bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgements}
I would like to thank my family for their support:
my wife Chao, my parents Norbert and Felizitas, and my sister Miriam.
Thanks to my office mates Oliver and Myles as well; it was always fun with you.
And finally, I thank my supervisors Novi and David.

\endgroup